AttCRISPR : a spacetime interpretable model for prediction of sgRNA on-target activity

### 2020-11-11 Paper/main.tex
11-11 paper

\title[short Title]{AttCRISPR : a spacetime interpretable model for prediction of sgRNA on-target activity}
\author[Xiao \textit{et~al}.]{Liming Xiao\,$^{\text{\sfb 1}}$, Yunqi Wan\,$^{\text{\sfb 1}}$ and Zhenran Jiang\,$^{\text{\sfb 1,}*}$}
\address{$^{\text{\sf 1}}$Department of Computer Science \& Technology, East China Normal University, Shanghai, 200062, China.}

@@

\abstract{\textbf{Motivation:} More and more Cas9 variants with higher specificities are developed to avoid the off-target effect, which bring a significant volume of experimental data. 
Conventional machine learning performs poorly on these datasets, while model based on deep learning are often lack of interpretability, which makes it difficult for researchers to understand its decisions, 
not to mention to design high-activity sgRNA with the insights brought by the model. 
Moreover, neither the deep learning based model with existing structure can not satisfy enough precision at such huge datasets.\\
\textbf{Results:} To overcome these, we design and implement AttCRISPR, a intrinsic interpretable method based on deep learning to predict the on-target activity. 
In this paper, we find that based on the preprocessing of the input base sequence, the deep learning models can be divided into two categories, 
methods in spatial domain with the input of encoded base sequence and in temporal domain with the input of embedded base sequence. 
Based on the categorization above, we assume that spatial and temporal method are heterogeneous, and use a full connection layer to stack them, which achieve a better performance. 
Our model was trained and tested on the biggest dataset, the same dataset as DeepHF used, as far as we know for performance evaluation. 
In addition, benefits from two attention modules, spatial and temporal attention modules, AttCRISPR has strong interpretability. 
Through them, we can understand the decisions made by AttCRISPR at both global and local levels without other post hoc explanations techniques. 
Further, interpretability AttCRISPR can be used to guide researcher to design sgRNA with higher activity, in order to reduce the experimental cost. 
AttCRISPR can compete with state-of-the-art method in these datasets, and achieves an average spearman value of 0.872, 0.867, 0.867 (corresponding to WT-SpCas9, eSpCas9(1.1), SpCas9-HF1) under tenfold shuffled validation. \\
\textbf{Availability:} The example code are available at \href{https://github.com/South-Walker/AttCRISPR}{https://github.com/South-Walker/AttCRISPR}\\
\textbf{Contact:} \href{zrjiang@cs.ecnu.edu.cn}{zrjiang@cs.ecnu.edu.cn}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

@@

In practice, there have been a number of application and toolkit applied in this task. In the earlier studies, methods in silico are categorized into three types: 
(1) alignment-based, (2) hypothesis-driven and (3) learning-based \citep{chuai2018deepcrispr}. 
Recently we noticed that the last type of method seems to be getting more attention because of huger and huger datasets \citep{liu2019computational}.

Learning-based method, which designed to predict the on-target activity (or off-target probability) of sgRNAs, 
is essentially a computational model built by machine learning algorithm, not only conventional machine learning but also deep learning algorithm. 
In general, the single or multiple base sequences (vary according to the task) and biological features are represented as a multi-dimensional vector $X\in\mathbb{R}^d$, 
and $d=l+b$, where $l$ refers to the length of base sequences and $b$ refers to the number of biological features, these methods can be represented as
\begin{equation}
 y=Score( X)\label{eq:01}
\end{equation}
 where $Score(\cdot)$ depends on the algorithm being selected, and $y$ denotes the predicted value.
 Some studies on HT\_ABE and HT\_CBE (two gene editing tools grew out of CRISPR) have shown that deep learning based models often outperformed conventional machine learning, when the number of sgRNAs in the dataset reached a certain level \citep{Song2020,kim2018deep,kim2019spcas9}. 
 Per contra, conventional machine learning algorithms, such as linear regression, logistic regression and the decision tree, are often more interpretable due to the fewer parameters and clearer mathematical assumptions. 
 In short, what was needed for developer is to trade-off accuracy and interpretability. 
 \citeauthor{MuhammadRafid2020} consider deep-learning models as black boxs and believe they lack interpretability, 
 motivated by the empirical assertion, they turn to build a model based conventional machine learning to compete with state-of-the-art deep learning models \citep{MuhammadRafid2020}. 
 On the other hand, input perturbation based feature importance analysis become a preferred components to reveal the importance of features in deep learning models. 
 \citeauthor{Liu2019} use a sliding window of length 2 to extract dimeric as input and rank the position of dimeric by contribution to final output \citep{Liu2019}. 
 One regret is that subject to the processing of the input sgRNA sequence, their analysis can not exactly on the nucleotide class. 
 Further, SHAP, one of the most prominent of model explain techniques, has been widely used to understand the decision made by the model. 
 \citeauthor{wang2019optimized} develope DeepHF, a deep learning based model, and use Deep SHAP to revealed nucleotide contributions \citep{wang2019optimized}. 
 Deep SHAP is a compositional approximation of SHAP values since it is challenged to computate SHAP values exactly, especially for a complex deep neural networks. 
 In our understanding, the method based on input perturbation often requires better generalization ability of the model (even for artificial ridiculous noise data).

 Beyond that, existing analysis of the models interpretability are all at global level, and result is a general pattern in the datasets. 
For example, which position has a great impact on the final decision of the model and which position-dependent nucleotide has a positive impact on the activity. 
Lack of analysis at local level. 
For example, which structure causes the low activity of a certain nucleotide sequence and how to improve its activity with a few modification. 
In light of the above, we believe it is essential to develope a model which can not only match deep learning based model in performance, 
but also with strong interpretability that can be comparable to conventional machine learning method.

@@

Deep neural network has shown its power in the study of CRISPR/Cas9 and its improved Systems \citep{liu2019computational}. 
 Most of the deep neural network existing are the combination of recurrent neural network (RNN), convolutional neural network (CNN), fully connected neural network (FNN), and their variants. 
 As Figure~\ref{fig:category} show,  
 We found that the deep learning models used in sgRNA on-target activity (even for off-target effect) prediction tasks in recent years can be divided into the following two categories according to the encoding approach of the sgRNA sequence (sgRNA-DNA sequence pair, for off-target effect prediction):

\begin{enumerate}
    \item Methods in spatial domain. 
    Some previous studies have used model based CNN to predict gRNA on-target activity or off-target effect \citep{lin2018off-target,kim2018deep,chuai2018deepcrispr}. 
    They process sgRNA base sequence inputs with the help of one-hot encoding idea. 
    In other words, they regard it as two-dimensional image data, and use convolution layer to extracte potential features in spatial domain, 
    It is worth noting that \citeauthor{zhang2020c-rnncrispr:} adds bidirectional gated recurrent unit (BGRU, in short), a RNN variant, after pooling layer of classic CNN network \citep{zhang2020c-rnncrispr:}. 
    Our explanation is that BGRU assists CNN to extract spatial features in one dimension, under this belief it belong to this category. 
    \item Methods in temporal domain. 
    Although RNN-based network have been shown effective to improve the performance of the model with temporal sequential input, especially in natural language processing and sequential recommendation, 
    RNN be not used for gRNA activity prediction, until recently \citep{wang2019optimized,liu2020deep,Liu2019}. 
    They consider the nucleotides(can also dimer or polymer) in the sgRNA sequence as word, and the sgRNA sequence itself as a sentence (from 5' to 3'), then a trainable matrix (could be either supervised or unsupervised) is used to project the word to the dense real-valued space. 
    This technology is called embedding, which generates the base embedding. RNN further encoding the base embedding into a sequence of hidden state vector. 
    Specially, \citeauthor{Liu2019} use RNN and CNN in parallel to extract features in base embedding \citep{Liu2019}. 
    However, base embedding is not spatially interpretable (different from one-hot encode), and they have no way to further explore the correlation between CNN and RNN output. 
    Almost all RNN based models used in sgRNA on-target activity or off-target effect flatten the hidden state vector into a one-dimensional vector as the input of the fully connected layer. 
    It is a pity that the temporal sequential dependency of hidden state vector are rarely noticed. 
    To summarise, RNN has limited representation power in capturing spatial feature. Furthermore, hidden state vector representation is usually hard to understand and explain.
\end{enumerate}

@@

\subsubsection{Method and attention in spatial domain}

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{CNNv2.png}}
    \caption{The architecture of spatial domain method in AttCRISPR. 
    The input of the method is encoded sgRNA sequence $X_{oh}$, a $21\times 4$ one-hot matrix. 
    Then refine it through a spatial attention module, which could tell us the importance of a specific matrix element (or just say, pixel).
    A simple convolutional neural network followed is applied to extract potential feature representation of sgRNA sequence.
    In the last step, we flatten the output of convolutional neural network into a one dimensional vector and 
    use a multilayer perceptron with sigmoid activation function to achieve the spatial output $y_s$.}\label{fig:CNN}
\end{figure}
As Figure~\ref{fig:CNN} show, method in spatial domain relies on the convolutional neural network. 
As previously mentioned, sgRNA sequence has been encoded into a $21\times 4$ one-hot matrix $X_{oh}$, and we regard $X_{oh}$ as a binary image. 
Then, convolution kernels with different size are used to extract potential spatial features just like what other do in computer vision. 
According to the foregoing, the Spatial Attention Module proposed by \citeauthor{woo2018cbam:} can be applied in our method in spatial domain, 
which was used to improve the performance of CNN in vision tasks.

@@

As Figure~\ref{fig:spatialmodule} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.

\subsubsection{Method and attention in temporal domain}
As Figure~\ref{fig:04} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy Equation~(\ref{eq:05}).

On this basis, we apply the idea of attention mechanism which has been widely used in natural language processing tasks to AttCRISPR in the method of temporal domain, 
and name it Temporal Attention Module \citep{vaswani2017attention,luong2015effective}.
As the idea of \citeauthor{vaswani2017attention}, Temporal Attention Module satisfies the following equation
\begin{equation}
Attention(Q,K,V)=align(Q,K)V\label{eq:13}
\end{equation}
where the $align(\cdot)$ is put forward by \citeauthor{luong2015effective}. 
$Q$, $K$, $V$ are queries, keys and values matrix accordingly in the paper of \citeauthor{vaswani2017attention}.

@@

\subsection{Experiments}

@@

The second experiment is designed to compare the performance of AttCRISPR and other current prediction method. 
In order to make the comparison apples to apples, we reduce the dimensionality of the same hand-crafted biological features as DeepHF's, 
which has been shown to enhance the predictability of a deep-learning model greatly, with a multilayer perceptron. 
Then follow the Equation~(\ref{eq:20}) to achieve the final prediction value. 
AttCRISPR (with the hand-crafted biological features) performs better on all three datasets than DeepHF, the current state-of-the-art.

@@

Our baselines have a comprehensive coverage of the methods tested in these datasets. 
In Table~\ref{Tab:baseline}, we annotate some properties of these baselines (is/isn't neural models, is/isn't end2end models)
All the experiments were carried out in Python 3.6 using Keras 2.2.4 and one GeForce RTX 2080Ti Super was used for training and testing if needed.

\section{Results}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (d)-(f) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

In Table~\ref{Tab:withoutbiofeature}, we list the performance of methods in spatial or temporal domain and the stacking of methods. 
Temporal AttCRISPR, TAC for short, achieved Spearman correlation coefficients of 0.857, 0.844, 0.851 in above three dataset. 
Spatial AttCRISPR, SpAC for short, corresponds to 0.862, 0.854, 0.857. In the absence of hand-crafted biological features. 
Ensemble AttCRISPR achieve the best performance of our knowledge, corresponds to 0.868, 0.859, 0.862.

In addition, in Table~\ref{Tab:withoutbiofeature}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Figure~\ref{fig:06} a-c). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

>>>>

\title[sgRNA on-target activity]{AttCRISPR : a spacetime interpretable model for prediction of sgRNA on-target activity}
\author[Xiao \textit{et~al}.]{Liming Xiao\,$^{\text{\sfb 1}}$, Yunqi Wan\,$^{\text{\sfb 1}}$ and Zhenran Jiang\,$^{\text{\sfb 1,}*}$}
\address{$^{\text{\sf 1}}$School of Computer Science \& Technology, East China Normal University, Shanghai, 200062, China.}

@@

\abstract{\textbf{Motivation:} More and more Cas9 variants with higher specificities are developed to avoid the off-target effect, which bring a significant volume of experimental data. 
Conventional machine learning performs poorly on these datasets, while the models based on deep learning are often lack of interpretability, which makes it difficult for researchers to understand its decisions, 
not to mention to design high-activity sgRNA with the insights brought by the model. 
Moreover, neither of the deep learning based model with existing structure can not satisfy enough precision at such huge datasets.\\
\textbf{Results:} To overcome these problems, we design and implement AttCRISPR, an intrinsic interpretable method based on deep learning to predict the on-target activity. 
In this paper, we find that based on the preprocessing of the input base sequence, the deep learning models can be divided into two categories, 
methods in spatial domain with the input of encoded base sequence and in temporal domain with the input of embedded base sequence. 
Based on the categorization above, we assume that spatial and temporal method are heterogeneous, and use a full connection layer to stack them, which achieve a better performance. 
Our model was trained and tested on the biggest dataset, the same dataset as DeepHF used, as far as we know for performance evaluation. 
In addition, benefits from two attention modules, spatial and temporal attention modules, AttCRISPR has strong interpretability. 
Through them, we can understand the decisions made by AttCRISPR at both global and local levels without other post hoc explanations techniques. 
Further, interpretability AttCRISPR can be used to guide researcher to design sgRNA with higher activity, in order to reduce the experimental cost. 
AttCRISPR can compete with state-of-the-art method in these datasets, and achieves an average spearman value of 0.872, 0.867, 0.867 (corresponding to WT-SpCas9, eSpCas9(1.1), SpCas9-HF1) under tenfold shuffled validation. \\
\textbf{Availability:} The example code are available at \href{https://github.com/South-Walker/AttCRISPR}{https://github.com/South-Walker/AttCRISPR}\\
\textbf{Contact:} \href{zrjiang@cs.ecnu.edu.cn}{zrjiang@cs.ecnu.edu.cn}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

@@

In practice, there have been a number of application and toolkits applied in this task. In the earlier studies, the methods \emph{in silico} are categorized into three types: 
(1) alignment-based, (2) hypothesis-driven and (3) learning-based \citep{chuai2018deepcrispr}. 
Recently, we noticed that the last type of method seems to be getting more attention because of huger and huger datasets \citep{liu2019computational}.

Learning-based method, which designed to predict the on-target activity (or off-target probability) of sgRNAs, 
is essentially a computational model built by machine learning algorithms, not only conventional machine learning but also deep learning. 
In general, the single or multiple base sequences (vary according to the task) and biological features are represented as a multi-dimensional vector $X\in\mathbb{R}^d$, 
and $d=l+b$, where $l$ refers to the length of base sequences and $b$ refers to the number of biological features, these methods can be represented as
\begin{equation}
 y=Score( X)\label{eq:01}
\end{equation}
 where $Score(\cdot)$ depends on the algorithm being selected, and $y$ denotes the predicted value.
 Some studies on HT\_ABE and HT\_CBE (two gene editing tools grew out of CRISPR) have shown that deep learning based models often outperformed conventional machine learning, when the number of sgRNAs in the dataset reached a certain level \citep{Song2020,kim2018deep,kim2019spcas9}. 
 Per contra, conventional machine learning algorithms, such as linear regression, logistic regression and the decision tree, are often more interpretable due to the fewer parameters and clearer mathematical assumptions. 
 In short, what was needed for developer is to trade-off accuracy and interpretability. 
 \citeauthor{MuhammadRafid2020} consider deep-learning models as black boxs and believe they lack interpretability, 
 motivated by the empirical assertion, they turn to build a model based conventional machine learning to compete with state-of-the-art deep learning models \citep{MuhammadRafid2020}. 
 On the other hand, input perturbation based feature importance analysis become a preferred components to reveal the importance of features in deep learning models. 
 \citeauthor{Liu2019} use a sliding window of length 2 to extract dimeric as input and rank the position of dimeric by contribution to final output \citep{Liu2019}. 
 One regret is that subject to the processing of the input sgRNA sequence, their analysis can not exactly on the nucleotide class. 
 Further, SHAP, one of the most prominent of model explain techniques, has been widely used to understand the decision made by the model. 
 \citeauthor{wang2019optimized} develope DeepHF, a deep learning based model, and use Deep SHAP to revealed nucleotide contributions \citep{wang2019optimized}. 
 Deep SHAP is a compositional approximation of SHAP values since it is challenged to computate SHAP values exactly, especially for a complex deep neural networks. 
 In our understanding, the method based on input perturbation often requires better generalization ability of the model (even for artificial ridiculous noise data).

 In addition, existing analysis of the models interpretability are all at global level, and result is a general pattern in the datasets. 
For example, which position has a great impact on the final decision of the model and which position-dependent nucleotide has a positive impact on the activity. 
Lack of analysis at local level. 
For example, which structure causes the low activity of a certain nucleotide sequence and how to improve its activity with a few modification. 
In light of the above, we believe it is essential to develope a model which can not only match deep learning based model in performance, 
but also with strong interpretability that can be comparable to conventional machine learning method.

@@

Deep neural network has shown its power in the study of CRISPR/Cas9 and its improved systems \citep{liu2019computational}. 
 Most of the deep neural networks existing are the combination of recurrent neural network (RNN), convolutional neural network (CNN), fully connected neural network (FNN), and their variants. 
 As Figure~\ref{fig:category} show,  
 We found that the deep learning models used in sgRNA on-target activity (even for off-target effect) prediction tasks in recent years can be divided into the following two categories according to the encoding approach of the sgRNA sequence (sgRNA-DNA sequence pair, for off-target effect prediction):

\begin{enumerate}
    \item Methods in spatial domain. 
    Some previous studies have used model based CNN to predict sgRNA on-target activity or off-target effect \citep{lin2018off-target,kim2018deep,chuai2018deepcrispr}. 
    They process sgRNA base sequence inputs with the help of one-hot encoding idea. 
    In other words, they regard it as two-dimensional image data, and use convolution layer to extracte potential features in spatial domain, 
    It is worth noting that \citeauthor{zhang2020c-rnncrispr:} adds bidirectional gated recurrent unit (BGRU, in short), a RNN variant, after pooling layer of classic CNN network \citep{zhang2020c-rnncrispr:}. 
    Our explanation is that BGRU assists CNN to extract spatial features in one dimension, under this belief it belong to this category. 
    \item Methods in temporal domain. 
    Although RNN-based network have been shown effective to improve the performance of the model with temporal sequential input, especially in natural language processing and sequential recommendation, 
    RNN be not used for gRNA activity prediction, until recently \citep{wang2019optimized,liu2020deep,Liu2019}. 
    They consider the nucleotides(can also dimer or polymer) in the sgRNA sequence as word, and the sgRNA sequence itself as a sentence (from 5' to 3'), then a trainable matrix (could be either supervised or unsupervised) is used to project the word to the dense real-valued space. 
    This technology is called embedding, which generates the base embedding. RNN further encoding the base embedding into a sequence of hidden state vector. 
    Specially, \citeauthor{Liu2019} use RNN and CNN in parallel to extract features in base embedding \citep{Liu2019}. 
    However, base embedding is not spatially interpretable (different from one-hot encode), and they have no way to further explore the correlation between CNN and RNN output. 
    Almost all RNN based models used in sgRNA on-target activity or off-target effect flatten the hidden state vector into a one-dimensional vector as the input of the fully connected layer. 
    It is a pity that the temporal sequential dependency of hidden state vector are rarely noticed. 
    To summarise, RNN has limited representation power in capturing spatial feature. Furthermore, hidden state vector representation is usually hard to understand and explain.
\end{enumerate}

@@

\subsubsection{Method in spatial domain}

\begin{figure}[!tpb]%figure2
    \centerline{\includegraphics[width=86mm]{CNNv2.png}}
    \caption{The architecture of spatial domain method in AttCRISPR. 
    The input of the method is encoded sgRNA sequence $X_{oh}$, a $21\times 4$ one-hot matrix. 
    Then refine it through a spatial attention module, which could tell us the importance of a specific matrix element (or just say, pixel).
    A simple convolutional neural network followed is applied to extract potential feature representation of sgRNA sequence.
    In the last step, we flatten the output of convolutional neural network into a one dimensional vector and 
    use a multilayer perceptron with sigmoid activation function to achieve the spatial output $y_s$.}\label{fig:CNN}
\end{figure}
As Figure~\ref{fig:CNN} demonstrated, the method in spatial domain relies on the convolutional neural network. 
As previously mentioned, sgRNA sequence has been encoded into a $21\times 4$ one-hot matrix $X_{oh}$, and we regard $X_{oh}$ as a binary image. 
Then, convolution kernels with different size are used to extract potential spatial features just like what other do in computer vision. 
According to the foregoing, the spatial attention Module proposed by \citeauthor{woo2018cbam:} can be applied in our method, 
which has been used to improve the performance of CNN in vision tasks.

@@

As shown in Figure~\ref{fig:spatialmodule}, given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.

\subsubsection{Method in temporal domain}
As Figure~\ref{fig:04} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy Equation~(\ref{eq:05}).

On this basis, we apply the idea of attention mechanism which has been widely used in natural language processing tasks to AttCRISPR in the method of temporal domain, 
and name it temporal attention module \citep{vaswani2017attention,luong2015effective}.
As the idea of \citeauthor{vaswani2017attention}, temporal attention module satisfies the following equation
\begin{equation}
Attention(Q,K,V)=align(Q,K)V\label{eq:13}
\end{equation}
where the $align(\cdot)$ is put forward by \citeauthor{luong2015effective}. 
$Q$, $K$, $V$ are queries, keys and values matrix accordingly in the paper of \citeauthor{vaswani2017attention}.

@@

\subsection{Experiment design}

@@

The second experiment is designed to compare the performance of AttCRISPR and other current prediction method. 
In order to make the comparison apples to apples, we reduce the dimensionality of the same hand-crafted biological features as DeepHF's, 
which has been shown to enhance the predictability of a deep-learning model greatly, with a multilayer perceptron. 
Then follow the Equation~(\ref{eq:20}) to achieve the final prediction value. 
AttCRISPR (with the hand-crafted biological features) performs better on all three datasets than DeepHF.

@@

Our baselines have a comprehensive coverage of the methods tested in these datasets. 
In Table~\ref{Tab:baseline}, we annotate some properties of these baselines (is/isn't neural models, is/isn't end2end models)
All of the experiments were carried out in Python 3.6 using Keras 2.2.4 and one GeForce RTX 2080Ti Super was used for training and testing if needed.

\section{Results}
\begin{figure}[!tpb]
    \centerline{\includegraphics[width=86mm]{baselinewithoutbiofeat.eps}}
    \caption{In the absence of hand-crafted biological features, performance of different algorithms for sgRNA activity prediction. 
    (a)-(c) The performance of Temporal AttCRISPR, Spatial AttCRISPR and Ensemble AttCRISPR. 
    The half-violin plots show the mean and distribution of the Spearman correlation coefficient between predicted and measured sgRNA activity scores over all tests. 
    (d)-(f) In the absence of hand-crafted biological features, the performance of all prediction methods in these datasets as far as we know. 
    The $mean \pm s.d.$ of the Spearman correlation coefficient between predicted and measured sgRNA activity scores are shown in the bar plots.}\label{fig:06}
\end{figure}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions made by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

In Table~\ref{Tab:withoutbiofeature}, we list the performance of methods in spatial or temporal domain and the stacking of methods. 
Temporal AttCRISPR, TAC for short, achieved Spearman correlation coefficients of 0.857, 0.844, 0.851 respectively in above three dataset. 
Spatial AttCRISPR, SpAC for short, corresponds to 0.862, 0.854, 0.857. In the absence of hand-crafted biological features. 
Ensemble AttCRISPR achieve the best performance of our knowledge, corresponds to 0.868, 0.859, 0.862.

In addition, in Table~\ref{Tab:withoutbiofeature}, the performance of other methods without using hand-crafted biological features, are also recorded. 
Regardless of the method we developed, RNN reported by \citeauthor{wang2019optimized}, which can be categorized as the method in temporal domain, 
is the most predictive with Spearman correlation coefficients of 0.856, 0.849, 0.851. 
It's obvious that the ensemble AttCRISPR is better at prediction (Figure~\ref{fig:06} (a-c)). 
Furthermore, the prediction ability of models could be boosted by addition of other hand-crafted biological features, which can't be obtained directly by sequence information.

Further experiment is designed to compare the performance of standard AttCRISPR (hand-crafted biological features are used to improve performance of ensemble AttCRISPR)
 and DeepHF, which is current state-of-the-art method.
\begin{table}[!tpb]
    \processtable{
    Performance comparisons for different methods in the absence of hand-crafted biological features 
    (take Spearman correlation coefficient as evaluation index)\label{Tab:withoutbiofeature}} 
    {\begin{tabular}{@{}lccc@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost$^*$ & 0.845 & 0.831 & 0.818\\
        MLP$^*$ & 0.842 & 0.846 & 0.844\\
        CNN$^*$ & 0.846 & 0.831 & 0.834\\
        RNN$^*$ & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        EnAC & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ)$^{\#}$ & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}{}
    \processtable{
    Performance comparisons for methods before and after integrating with hand-crafted biological features
(take Spearman correlation coefficient as evaluation index)\label{Tab:withbiofeature}} 
    {\begin{tabular}{lcccccc}\toprule
        & \multicolumn{2}{c}{WT-SpCas9} & \multicolumn{2}{c}{eSpCas9(1.1)} & \multicolumn{2}{c}{SpCas9-HF} \\ 
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \noalign{\smallskip} 
        Method & Mean  & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} & Mean & \begin{tabular}[c]{@{}l@{}}Std Dev\\ ($\times 10^{-3}$)\end{tabular} \\ \hline
        RNN$^*$ & 0.856 & 3.33 & 0.849 & 5.00 & 0.851 & 4.11 \\
        EnAC   & 0.868 & 2.66 & 0.859 & 4.66 & 0.862 & 3.19 \\
        DeepHF$^*$ & 0.867 & \textbf{2.37} & 0.862 & 4.24 & 0.860 & 3.21 \\
        StAC   & \textbf{0.872} & 2.55 & \textbf{0.867} & \textbf{3.71} & \textbf{0.867} & \textbf{2.65} \\
     \botrule
\end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: 
The method with superscript of $^*$ and $^{\#}$ is reported by \citeauthor{wang2019optimized} and \citeauthor{MuhammadRafid2020} respectively. 
In the tables, we use the results reported in the relevant papers as the performance of the method directly.}
\end{table}

\subsection{Performance comparison with current methods}\label{section:comparison}

In order to make the comparison apples to apples, 
we follow Equation~(\ref{eq:20}) to modify the ensemble method and design the control experiment using the same strategy to validate the conclusion that integrating with hand-crafted biological features can improve the predictive performance of methods. 
What's more, we compare the standard AttCRISPR and DeepHF (Table~\ref{Tab:withbiofeature}).
